### 2013-03-28 doc/sosp13/evaluation.tex
benchmark plot
macrotize all of grappa

\paragraph{Global memory and communication.} We measure the performance of
\Grappa's global memory and communication layers using a faithful
implementation of the giga updates per second (GUPS) benchmark.
Read-modify-write updates are dispatched at random to a large array. This
benchmark stresses the communication layer of \Grappa separately from the
scheduler, because only a single worker is used per system node. Figure~\ref{fig:grappa-gups} shows that Grappa is able to sustain on the order of billions of updates per second, which is enough to \TODO{...}.

@@

\begin{figure}[ht]
    \begin{center}
%      \includegraphics[width=0.5\textwidth]{figs/grappa-uts.pdf}
    \end{center}
    \caption{Comparing \Grappa with XMT, hand-tuned MPI and UPC.}
%bar chart, one set of bars per benchmark, one bar per system. runtime normalized to grappa.
    \label{fig:grappa-comparisons}
\end{figure}

@@

Our evaluation begins with presentation of microbenchmark results, establishing
the intrinsic potential of Grappa to provide random access bandwidth and latency tolerance. 
Next, we present application results, both for Grappa and for other paradigms, as well
as comparing against the Cray XMT.  Finally, we present the impact of increased aggregation delay
on Grappa results, thus exploring robustness to network scale.

@@

\subsubsection{Random Access}
\TODO{Random access feed forward results on Grappa.  optional: Results we measured for XMT.  Cite MPI results}
\subsubsection{Latency Tolerance}
\TODO{Simple ping test results -- eg, MPI ping, not the full blown aggregation ping test.  Random access blocking results on Grappa.}
\subsubsection{Scheduling and Robustness}

@@

\subsection{Application Results}
To evaluate Grappa's performance with respect to the XMT, we ran each
of our three benchmarks on up to 16 nodes of each machine. Grappa used
6 cores per node, with the best parameters chosen for each point. In
some cases, the XMT could not run the benchmark with 2 nodes, so the
point is omitted. \TODO{rewrite this!}
\subsubsection{Unbalanced Tree Search}\TODO{rewrite with new results}
%% UTS: performance comparison
\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/uts_performance.pdf}
    \end{center}
    \caption{Performance of in-memory unbalanced tree search.}
    \label{fig:uts_compare}
\end{figure}

We ran UTS-mem with a geometric 100M-vertex tree
(T1L). Figure~\ref{fig:uts_compare} shows the performance in terms of
number of vertices visited per second versus number of compute
nodes. Grappa is 3.2 times faster than the XMT at 16 nodes.  As we will show later, the performance advantage Grappa has over XMT increases as more nodes are added.  The main reason Grappa performs better is the software-based delegate synchronization obviates the need for the retry-based synchronization that XMT uses.

@@

We ran BFS on a synthetic Kronecker graph with $2^{25}$ vertices and
$2^{29}$ edges (25 GB of data). Figure~\ref{fig:bfs-performance} shows
our performance in terms of graph edges traversed per second. The XMT
is 2.5 times faster than Grappa at 16 nodes.  Performance does scale at a constant rate for Grappa, suggesting that adding more nodes will increase performance.

@@

We ran Betweenness Centrality on the same scale 25 Kronecker graph as
we did for BFS. Figure~\ref{fig:centrality-performance} shows our
performance in terms of graph edges traversed per second. At 16 XMT
processors/cluster nodes, the XMT is 1.75 times faster than Grappa.

% Got rid of this discussion of UTS by itself, tried to work the highlights in below
%\paragraph{UTS-Mem}
%We ran UTS-Mem on Grappa and the XMT with a geometric 1.6B-vertex tree
%(T1XL) and a geometric 4.2B-vertex tree (T1XXL), using up to 128
%nodes---the maximum we had available for each. Grappa results are for 5 cores per node. Grappa with 20 machines is faster than the entire XMT of 128 processors.
%Grappa achieves \checkme{188Mvert/s} with 128 nodes and the XMT
%achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes, Grappa adds 1.4 Mvert/s/node.
%The XMT scales at 850 Kvert/s/node, until it plateaus. Grappa keeps
%scaling up through 128 nodes, although scaling
%declines because of the unscalability of our aggregation mechanism as
%number of network endpoints increases. 
%
%Despite our efforts to tune the UTS implementation specific to the 
%XMT, performance does not scale well with increasing processor count,
%flattening out around 60 processors.  When we increase the size of
%the tree from 100M to 4.2B, we find that performance does not improve,
%suggesting that performance is not limited by task parallelism.
%Cray's performance tools show an increasing number of memory
%retry operations for failed synchronization operations generated by
%the runtime, which create network contention.
%
To determine how Grappa's performance scales compared to the performance of the entire XMT, we ran a set of experiments up to all 128 XMT processors and 128 cluster nodes. For the XMT, the number of allowed processors was varied up to the entire machine, with some minor tuning of stream parameters needed to get optimal performance. For Grappa, parameters such as cores per node, aggregator timeouts, and parallel threshold were tuned to get the best performance for each node count. All of the benchmarks continue to improve out to 128 nodes for Grappa. UTS continues to fare better than the XMT with large node counts, with the XMT appearing to plateau at 60 processors due to contention from synchronization retries, while Grappa handles this by suspending tasks until messages return. For BFS and Centrality, the XMT scales approximately a constant factor better than Grappa. We attribute this to a limitation in the current aggregator design and network stack that Grappa uses.  This limits the practical number of cores we can use to 6 per node (adding more cores per node \emph{decreases} performance).  Ironically, this limitation makes Grappa applications compute-bound instead of network-bound.  Work is ongoing to rework the Infiniband driver stack and aggregation interface to remove this limitation and improve aggregation addressing using local routing.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/scaling_cropped.pdf}
    \end{center}
    \caption{Scaling number of nodes: Grappa continues to perform significantly better than XMT for UTS but scales a constant factor slower than XMT for BFS (4x slower) and Centrality (2x slower). }
    \label{fig:uts_threshold}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/aggregator_ping}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:aggregator-ping} Bandwidth versus message size
    unidirectional ping test for Grappa with aggregation, Grappa with
    raw GASNet messages, and MPI. Aggregation provides an 11x
    bandwidth benefit at our common operating point.}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

@@

To implement the ping test, we wrote a simple Grappa application where
the cores of one node send messages as fast as possible to the cores
of another node. We vary the size of the payload up the maximum
payload size supported by the aggregator (nearly 4KB). Each core has a
single task sending to a single destination, so this is a best case
scenario for the aggregator. To see the benefit of the aggregator, we
added a bypass that lets us send messages directly through GASNet. We
also compare against the OSU \texttt{osu\_mbw\_mr} benchmark
\cite{osu:mpi}  compiled against OpenMPI 1.5.3; this
benchmark has the same pattern of communication but doesn't have the
overhead of Grappa's context switching.

@@

First, small message performance against the existing libraries is, as expected, poor. The MPI application test shows us that peak per node
bandwidth supported by our infiniband card is 3.4GB/s. This is
achievable only with large messages; we must send 16KB packets to get
within 5 percent of peak bandwidth. But in our benchmarks, we saw
average message between 32 and 64 bytes. At 32 bytes, the MPI test is
using less than 7 percent of its peak bandwidth. Grappa sending
messages directly through GASNet uses less than 3 percent of the peak
bandwidth.

Second, aggregation has the potential to improve this situation by an
order of magnitude. With aggregation, Grappa is able to send 32-byte
messages over 12 times faster than using GASNet directly. This is a
more respectable 32 percent of peak bandwidth. Due to expedient design
decisions, Grappa's aggregator limits its aggregation to 4KB; this
limits its peak achievable bandwidth to 75 percent of the actual
peak.

@@

\begin{figure}[htb]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/no_aggregation_uts.pdf}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:no-aggregation-uts} Performance of UTS on 16
      nodes with and without Grappa's aggregation.}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

\subsection{Scaling}
% Got rid of this discussion of UTS by itself, tried to work the highlights in below
%\paragraph{UTS-Mem}
%We ran UTS-Mem on Grappa and the XMT with a geometric 1.6B-vertex tree
%(T1XL) and a geometric 4.2B-vertex tree (T1XXL), using up to 128
%nodes---the maximum we had available for each. Grappa results are for 5 cores per node. Grappa with 20 machines is faster than the entire XMT of 128 processors.
%Grappa achieves \checkme{188Mvert/s} with 128 nodes and the XMT
%achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes, Grappa adds 1.4 Mvert/s/node.
%The XMT scales at 850 Kvert/s/node, until it plateaus. Grappa keeps
%scaling up through 128 nodes, although scaling
%declines because of the unscalability of our aggregation mechanism as
%number of network endpoints increases. 
%
%Despite our efforts to tune the UTS implementation specific to the 
%XMT, performance does not scale well with increasing processor count,
%flattening out around 60 processors.  When we increase the size of
%the tree from 100M to 4.2B, we find that performance does not improve,
%suggesting that performance is not limited by task parallelism.
%Cray's performance tools show an increasing number of memory
%retry operations for failed synchronization operations generated by
%the runtime, which create network contention.
%
To determine how Grappa's performance scales compared to the performance of the entire XMT, we ran a set of experiments up to all 128 XMT processors and 128 cluster nodes. For the XMT, the number of allowed processors was varied up to the entire machine, with some minor tuning of stream parameters needed to get optimal performance. For Grappa, parameters such as cores per node, aggregator timeouts, and parallel threshold were tuned to get the best performance for each node count. All of the benchmarks continue to improve out to 128 nodes for Grappa. UTS continues to fare better than the XMT with large node counts, with the XMT appearing to plateau at 60 processors due to contention from synchronization retries, while Grappa handles this by suspending tasks until messages return. For BFS and Centrality, the XMT scales approximately a constant factor better than Grappa. We attribute this to a limitation in the current aggregator design and network stack that Grappa uses.  This limits the practical number of cores we can use to 6 per node (adding more cores per node \emph{decreases\/} performance).  Ironically, this limitation makes Grappa applications compute-bound instead of network-bound.  Work is ongoing to rework the Infiniband driver stack and aggregation interface to remove this limitation and improve aggregation addressing using local routing.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/scaling_cropped.pdf}
    \end{center}
    \caption{Scaling number of nodes: Grappa continues to perform significantly better than XMT for UTS but scales a constant factor slower than XMT for BFS (4x slower) and Centrality (2x slower). }
    \label{fig:uts_threshold}
\end{figure}

@@

Parallel overhead---in the form of context switches, task spawns, and
synchronization---can reduce the performance benefit of parallelism.
Grappa sees a benefit to limiting the amount of parallelism created by
a recursive loop decomposition. The parallel loop threshold (``parallel
granularity'') parameter tells the runtime when to stop creating new tasks and just execute iterations
sequentially. This allows us to amortize the overhead of task
creation. In addition, assigning sequential iterations to a single
task provides the potential to exploit locality when data for adjacent iterations is
also adjacent in memory. The ability to exploit this locality that
exists in the application is an important advantage. We found that in UTS and BFS, increasing the
threshold from 1 up to 8 or 16, respectively, increases performance by
more than 60\%.

@@

%Retries are performed by the memory controller when remote synchronization operations fail to find the full-bit associated with each memory location in the unavailable state.  Retries are issued at low priority relative to new memory operations issued by the processor by other contexts, so they consume what would otherwise be unused injection bandwidth.  On a full-bandwidth system such as the MTA-2, retries have no impact on the progress of tasks other than their own.  On a Cray XMT, network bandwidth is limited, so retries create congestion.  In comparison, Grappa performs synchronization without retries, delaying responses at the receiving end until ready to notify the sender to proceed.  This saves bandwidth and permits scaling of tasks performing synchronization even on low injection rate networks.

>>>>

\paragraph{Global memory and communication.} We measure the performance of
\Grappa's global memory and communication layers using a faithful
implementation of the giga updates per second (GUPS) benchmark.
Read-modify-write updates are dispatched at random to a large array. This
benchmark stresses the communication layer of \Grappa separately from the
scheduler, because only a single worker is used per system node. Figure~\ref{fig:grappa-gups} shows that \Grappa is able to sustain on the order of billions of updates per second, which is enough to \TODO{...}.

@@

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{results/benchmarks.pdf}
    \end{center}
    \caption{Comparing \Grappa with XMT, hand-tuned MPI and UPC.}
%bar chart, one set of bars per benchmark, one bar per system. runtime normalized to grappa.
    \label{fig:grappa-comparisons}
\end{figure}

@@

Our evaluation begins with presentation of microbenchmark results, establishing
the intrinsic potential of \Grappa to provide random access bandwidth and latency tolerance. 
Next, we present application results, both for \Grappa and for other paradigms, as well
as comparing against the Cray XMT.  Finally, we present the impact of increased aggregation delay
on \Grappa results, thus exploring robustness to network scale.

@@

\subsubsection{Random Access}
\TODO{Random access feed forward results on \Grappa.  optional: Results we measured for XMT.  Cite MPI results}
\subsubsection{Latency Tolerance}
\TODO{Simple ping test results -- eg, MPI ping, not the full blown aggregation ping test.  Random access blocking results on \Grappa.}
\subsubsection{Scheduling and Robustness}

@@

\subsection{Application Results}
To evaluate \Grappa's performance with respect to the XMT, we ran each
of our three benchmarks on up to 16 nodes of each machine. \Grappa used
6 cores per node, with the best parameters chosen for each point. In
some cases, the XMT could not run the benchmark with 2 nodes, so the
point is omitted. \TODO{rewrite this!}
\subsubsection{Unbalanced Tree Search}\TODO{rewrite with new results}
%% UTS: performance comparison
\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/uts_performance.pdf}
    \end{center}
    \caption{Performance of in-memory unbalanced tree search.}
    \label{fig:uts_compare}
\end{figure}

We ran UTS-mem with a geometric 100M-vertex tree
(T1L). Figure~\ref{fig:uts_compare} shows the performance in terms of
number of vertices visited per second versus number of compute
nodes. \Grappa is 3.2 times faster than the XMT at 16 nodes.  As we will show later, the performance advantage \Grappa has over XMT increases as more nodes are added.  The main reason \Grappa performs better is the software-based delegate synchronization obviates the need for the retry-based synchronization that XMT uses.

@@

We ran BFS on a synthetic Kronecker graph with $2^{25}$ vertices and
$2^{29}$ edges (25 GB of data). Figure~\ref{fig:bfs-performance} shows
our performance in terms of graph edges traversed per second. The XMT
is 2.5 times faster than \Grappa at 16 nodes.  Performance does scale at a constant rate for \Grappa, suggesting that adding more nodes will increase performance.

@@

We ran Betweenness Centrality on the same scale 25 Kronecker graph as
we did for BFS. Figure~\ref{fig:centrality-performance} shows our
performance in terms of graph edges traversed per second. At 16 XMT
processors/cluster nodes, the XMT is 1.75 times faster than \Grappa.

% Got rid of this discussion of UTS by itself, tried to work the highlights in below
%\paragraph{UTS-Mem}
%We ran UTS-Mem on \Grappa and the XMT with a geometric 1.6B-vertex tree
%(T1XL) and a geometric 4.2B-vertex tree (T1XXL), using up to 128
%nodes---the maximum we had available for each. \Grappa results are for 5 cores per node. \Grappa with 20 machines is faster than the entire XMT of 128 processors.

%\Grappa achieves \checkme{188Mvert/s} with 128 nodes and the XMT
%achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes, \Grappa adds 1.4 Mvert/s/node.
%The XMT scales at 850 Kvert/s/node, until it plateaus. \Grappa keeps
%scaling up through 128 nodes, although scaling
%declines because of the unscalability of our aggregation mechanism as
%number of network endpoints increases. 
%
%Despite our efforts to tune the UTS implementation specific to the 
%XMT, performance does not scale well with increasing processor count,
%flattening out around 60 processors.  When we increase the size of
%the tree from 100M to 4.2B, we find that performance does not improve,
%suggesting that performance is not limited by task parallelism.
%Cray's performance tools show an increasing number of memory
%retry operations for failed synchronization operations generated by
%the runtime, which create network contention.
%
To determine how \Grappa's performance scales compared to the performance of the entire XMT, we ran a set of experiments up to all 128 XMT processors and 128 cluster nodes. For the XMT, the number of allowed processors was varied up to the entire machine, with some minor tuning of stream parameters needed to get optimal performance. For \Grappa, parameters such as cores per node, aggregator timeouts, and parallel threshold were tuned to get the best performance for each node count. All of the benchmarks continue to improve out to 128 nodes for \Grappa. UTS continues to fare better than the XMT with large node counts, with the XMT appearing to plateau at 60 processors due to contention from synchronization retries, while \Grappa handles this by suspending tasks until messages return. For BFS and Centrality, the XMT scales approximately a constant factor better than \Grappa. We attribute this to a limitation in the current aggregator design and network stack that \Grappa uses.  This limits the practical number of cores we can use to 6 per node (adding more cores per node \emph{decreases} performance).  Ironically, this limitation makes \Grappa applications compute-bound instead of network-bound.  Work is ongoing to rework the Infiniband driver stack and aggregation interface to remove this limitation and improve aggregation addressing using local routing.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/scaling_cropped.pdf}
    \end{center}
    \caption{Scaling number of nodes: \Grappa continues to perform significantly better than XMT for UTS but scales a constant factor slower than XMT for BFS (4x slower) and Centrality (2x slower). }
    \label{fig:uts_threshold}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/aggregator_ping}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:aggregator-ping} Bandwidth versus message size
    unidirectional ping test for \Grappa with aggregation, \Grappa with
    raw GASNet messages, and MPI. Aggregation provides an 11x
    bandwidth benefit at our common operating point.}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

@@

To implement the ping test, we wrote a simple \Grappa application where
the cores of one node send messages as fast as possible to the cores
of another node. We vary the size of the payload up the maximum
payload size supported by the aggregator (nearly 4KB). Each core has a
single task sending to a single destination, so this is a best case
scenario for the aggregator. To see the benefit of the aggregator, we
added a bypass that lets us send messages directly through GASNet. We
also compare against the OSU \texttt{osu\_mbw\_mr} benchmark
\cite{osu:mpi}  compiled against OpenMPI 1.5.3; this
benchmark has the same pattern of communication but doesn't have the
overhead of \Grappa's context switching.

@@

First, small message performance against the existing libraries is, as expected, poor. The MPI application test shows us that peak per node
bandwidth supported by our infiniband card is 3.4GB/s. This is
achievable only with large messages; we must send 16KB packets to get
within 5 percent of peak bandwidth. But in our benchmarks, we saw
average message between 32 and 64 bytes. At 32 bytes, the MPI test is
using less than 7 percent of its peak bandwidth. \Grappa sending
messages directly through GASNet uses less than 3 percent of the peak
bandwidth.

Second, aggregation has the potential to improve this situation by an
order of magnitude. With aggregation, \Grappa is able to send 32-byte
messages over 12 times faster than using GASNet directly. This is a
more respectable 32 percent of peak bandwidth. Due to expedient design
decisions, \Grappa's aggregator limits its aggregation to 4KB; this
limits its peak achievable bandwidth to 75 percent of the actual
peak.

@@

\begin{figure}[htb]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/no_aggregation_uts.pdf}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:no-aggregation-uts} Performance of UTS on 16
      nodes with and without \Grappa's aggregation.}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

\subsection{Scaling}
% Got rid of this discussion of UTS by itself, tried to work the highlights in below
%\paragraph{UTS-Mem}
%We ran UTS-Mem on \Grappa and the XMT with a geometric 1.6B-vertex tree
%(T1XL) and a geometric 4.2B-vertex tree (T1XXL), using up to 128
%nodes---the maximum we had available for each. \Grappa results are for 5 cores per node. \Grappa with 20 machines is faster than the entire XMT of 128 processors.
%\Grappa achieves \checkme{188Mvert/s} with 128 nodes and the XMT
%achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes, \Grappa adds 1.4 Mvert/s/node.
%The XMT scales at 850 Kvert/s/node, until it plateaus. \Grappa keeps
%scaling up through 128 nodes, although scaling
%declines because of the unscalability of our aggregation mechanism as
%number of network endpoints increases. 
%
%Despite our efforts to tune the UTS implementation specific to the 
%XMT, performance does not scale well with increasing processor count,
%flattening out around 60 processors.  When we increase the size of
%the tree from 100M to 4.2B, we find that performance does not improve,
%suggesting that performance is not limited by task parallelism.
%Cray's performance tools show an increasing number of memory
%retry operations for failed synchronization operations generated by
%the runtime, which create network contention.
%
To determine how \Grappa's performance scales compared to the performance of the entire XMT, we ran a set of experiments up to all 128 XMT processors and 128 cluster nodes. For the XMT, the number of allowed processors was varied up to the entire machine, with some minor tuning of stream parameters needed to get optimal performance. For \Grappa, parameters such as cores per node, aggregator timeouts, and parallel threshold were tuned to get the best performance for each node count. All of the benchmarks continue to improve out to 128 nodes for \Grappa. UTS continues to fare better than the XMT with large node counts, with the XMT appearing to plateau at 60 processors due to contention from synchronization retries, while \Grappa handles this by suspending tasks until messages return. For BFS and Centrality, the XMT scales approximately a constant factor better than \Grappa. We attribute this to a limitation in the current aggregator design and network stack that \Grappa uses.  This limits the practical number of cores we can use to 6 per node (adding more cores per node \emph{decreases\/} performance).  Ironically, this limitation makes \Grappa applications compute-bound instead of network-bound.  Work is ongoing to rework the Infiniband driver stack and aggregation interface to remove this limitation and improve aggregation addressing using local routing.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/scaling_cropped.pdf}
    \end{center}
    \caption{Scaling number of nodes: \Grappa continues to perform significantly better than XMT for UTS but scales a constant factor slower than XMT for BFS (4x slower) and Centrality (2x slower). }
    \label{fig:uts_threshold}
\end{figure}

@@

Parallel overhead---in the form of context switches, task spawns, and
synchronization---can reduce the performance benefit of parallelism.
\Grappa sees a benefit to limiting the amount of parallelism created by
a recursive loop decomposition. The parallel loop threshold (``parallel
granularity'') parameter tells the runtime when to stop creating new tasks and just execute iterations
sequentially. This allows us to amortize the overhead of task
creation. In addition, assigning sequential iterations to a single
task provides the potential to exploit locality when data for adjacent iterations is
also adjacent in memory. The ability to exploit this locality that
exists in the application is an important advantage. We found that in UTS and BFS, increasing the
threshold from 1 up to 8 or 16, respectively, increases performance by
more than 60\%.

@@

%Retries are performed by the memory controller when remote synchronization operations fail to find the full-bit associated with each memory location in the unavailable state.  Retries are issued at low priority relative to new memory operations issued by the processor by other contexts, so they consume what would otherwise be unused injection bandwidth.  On a full-bandwidth system such as the MTA-2, retries have no impact on the progress of tasks other than their own.  On a Cray XMT, network bandwidth is limited, so retries create congestion.  In comparison, \Grappa performs synchronization without retries, delaying responses at the receiving end until ready to notify the sender to proceed.  This saves bandwidth and permits scaling of tasks performing synchronization even on low injection rate networks.

### 2013-03-28 doc/sosp13/intro.tex
macrotize all of grappa

Our evaluation of Grappa shows that it runs several irregular application
kernels (e.g., graph traversal and sort) very efficiently on a commodity
cluster. Our yardstick for comparison is the XMT hardware itself. Using the
same number of network interfaces, \Grappa is in the same ballpark as the XMT:
For unbalanced tree search, Grappa is over \TODO{3X} faster and shows greatly
improved scalability; conversely, for breadth first search and betweenness
centrality Grappa is 2.5X slower. In Section~\ref{sec:evaluation} we explore
the factors that underpin this performance. Most importantly, however, for
significantly less real world cost, users can \emph{add\/} significantly more
processors to a commodity cluster than an XMT machine and use Grappa to
achieve scalable performance. \TODO{revise paragraph after eval, including MPI
and possibly programming model}

>>>>

Our evaluation of \Grappa shows that it runs several irregular application
kernels (e.g., graph traversal and sort) very efficiently on a commodity
cluster. Our yardstick for comparison is the XMT hardware itself. Using the
same number of network interfaces, \Grappa is in the same ballpark as the XMT:
For unbalanced tree search, \Grappa is over \TODO{3X} faster and shows greatly
improved scalability; conversely, for breadth first search and betweenness
centrality \Grappa is 2.5X slower. In Section~\ref{sec:evaluation} we explore
the factors that underpin this performance. Most importantly, however, for
significantly less real world cost, users can \emph{add\/} significantly more
processors to a commodity cluster than an XMT machine and use \Grappa to
achieve scalable performance. \TODO{revise paragraph after eval, including MPI
and possibly programming model}
